A matrix-free approach for finite-strain hyperelastic problems using geometric multigrid

### 2019-01-31 doc/large_strain.tex
make GMG MPI picture consistent with fine level one (use 4 procs in both)
working table with MPI/SIMD speedups
plot 3D speedup table
cleanup MPI/SIMD tables
fixed precision=3 for tables
remove obsolete todo

@@

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.8\textwidth}
    \centering
    \includegraphics[width=\textwidth]{gmg_2d.png}
    % \caption{2D multigrid mesh with 2 global refinements}
  \end{subfigure}
  \caption{2D multigrid mesh after two global refinements distributed into three MPI processes. The color indicates the process owning the element.}%
  \label{fig:miehe_gmg}
\end{figure}
%
%
{\color{red}
MPI parallelization is achieved using the standard domain decomposition, where each MPI process ``owns'' only a subset of elements.
To that end, \texttt{deal.II}'s interface to \texttt{p4est} \cite{p4est} library is used.
Figure \ref{fig:miehe_fine_level} illustrates such decomposition on the fine mesh level for the 2D problem.
The same approach is adopted on the multigrid levels.}
Figure \ref{fig:miehe_gmg} illustrates the hierarchy of multigrid meshes for the 2D problem after two global refinements.
{\color{red}
Implementational details of the MPI-parallel multigrid method in \texttt{deal.II} are discussed in \cite{Clevenger2018}.
}

@@

As outlined in the introduction, it is not only the performance of the matrix-free vector multiplication, but also the memory requirement to store a sparse matrix which is the driving force behind matrix-free methods.
Figures \ref{fig:benchmark_miehe_Emmy_memory2}, \ref{fig:benchmark_miehe_Emmy_memory3}, \ref{fig:benchmark_miehe_IWR_memory2} and \ref{fig:benchmark_miehe_IWR_memory3} demonstrate that even with caching one fourth-order symmetric tensor and one second-order tensor at each quadrature point, the matrix-free approach takes much less memory than its matrix-based counterpart.
{\color{red}
In those graphs we report the memory needed to cache additional quantities required for the tangent operator here considered,
as well as the memory required by the \texttt{deal.II} \textit{MatrixFree} object, which caches the inverse of the Jacobian of the transformation
from unit to real cell \cite{kronbichler12}.
Note that for Algorithm \ref{alg:mf_scalar} we need to evaluate gradients with respect to both deform and undeformed
configurations, and therefore two inverses will be cached via two different \textit{MatrixFree} objects.\todo{rerun IWR}\todo{this is roughly main memory transfer, do we elaborate on this/use it?}}

@@

>>>>

\usepackage{pgfkeys,pgfmath,pgfcore}
% \pgfkeys{/pgf/number format/.cd,1000 sep={}}
\pgfkeys{/pgf/number format/fixed,/pgf/number format/precision=3}

@@

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{0.8\textwidth}
    \centering
    \includegraphics[width=\textwidth]{gmg_2d_4procs.png}
    % \caption{2D multigrid mesh with 2 global refinements}
  \end{subfigure}
  \caption{2D multigrid mesh after two global refinements distributed into four MPI processes. The color indicates the process owning the element.}%
  \label{fig:miehe_gmg}
\end{figure}
%
%
{\color{red}
MPI parallelization is achieved using the standard domain decomposition, where each MPI process ``owns'' only a subset of elements.
To that end, \texttt{deal.II}'s interface to \texttt{p4est} \cite{p4est} library is used.
Figure \ref{fig:miehe_fine_level} illustrates such decomposition on the fine mesh level for the 2D problem.
The same approach is adopted on the multigrid levels.}
Figure \ref{fig:miehe_gmg} illustrates the hierarchy of multigrid meshes for the 2D problem after two global refinements.
{\color{red}
Implementational details of the MPI-parallel multigrid method in \texttt{deal.II} are discussed in \cite{Clevenger2018}.
}

@@

As outlined in the introduction, it is not only the performance of the matrix-free vector multiplication, but also the memory requirement to store a sparse matrix which is the driving force behind matrix-free methods.
Figures \ref{fig:benchmark_miehe_Emmy_memory2}, \ref{fig:benchmark_miehe_Emmy_memory3}, \ref{fig:benchmark_miehe_IWR_memory2} and \ref{fig:benchmark_miehe_IWR_memory3} demonstrate that even with caching one fourth-order symmetric tensor and one second-order tensor at each quadrature point, the matrix-free approach takes much less memory than its matrix-based counterpart.
{\color{red}
In those graphs we report the memory needed to cache additional quantities required for the tangent operator here considered,
as well as the memory required by the \texttt{deal.II} \textit{MatrixFree} object, which caches the inverse of the Jacobian of the transformation
from unit to real cell \cite{kronbichler12}.
Note that for Algorithm \ref{alg:mf_scalar} we need to evaluate gradients with respect to both deform and undeformed
configurations, and therefore two inverses will be cached via two different \textit{MatrixFree} objects.\todo{this is roughly main memory transfer, do we elaborate on this/use it?}}

@@

\todo{Discuss MPI/SIMD}
\input{parallelization_2d.tex}
\input{parallelization_3d.tex}
